Multi-task learning in Computer Vision

### 2020-05-28 __latexindent_temp.tex
Add some conclusive remarks

\chapter{Neural networks}
The topics presented in this thesis are based on neural networks, especially convolutional neural networks, which are a distinct neural network variation for image processing using filters.
Here we provide a very brief high-level overview of some of the concepts that are important for the later chapters, but a more detailed overview of deep learning and neural networks can be found, for example, in \citep{deep-learning}.

\section{Convolutional Neural Networks}
The most significant component of Convolutional Neural Networks (CNNs) is the convolutional layers that are their primary building block.
Using convolutional filters allows neural networks to extract useful information from images to accomplish various tasks.
The convolutional blocks are based on the mathematical convolution operation of capturing the signal of pre-determined window size.
A CNN classifier often consists of multiple convolutional layers with other layers in between.
Successive convolutional layers build on top of one another, and the deeper layers get high activations on more specific features.
In many final models, the earlier layers are thus somewhat similar and re-usable as they often detect relatively primitive shapes, like diagonal or horizontal lines.
In this thesis, we often talk about the parameters or the weights of a network.
The number of parameters in a network tells how complex it is and how expensive it is to store.
The floating-point numbers in the various layers are these weights, which are generally 32-bit numbers, but their precision can vary.

\subsection{Parameters}
To get the total number of parameters required for a single convolutional layer, we need to determine the input size, convolution window size, and the number of filters.
For example, if the input is a 256x256 RGB image, the input size is 256x256x3.
Now, if we apply 16 filters with a window size of 5, each of the filters requires 5x5x3 parameters, bringing the total to 1200 parameters.
A fully connected layer with 1000 hidden units would require a matrix with around 200 million parameters, so it would not really be possible even to apply multiple of these layers while maintaining the image size.
The convolution allows for keeping the dimensions between successive layers and only modifying the number of channels between the layers.
Often convolutional models reduce the first two dimensions, and the number of channels increases in the deeper layers.
For this, the convolutions can use a stride parameter to skip some positions for the convolution.
The pooling layer is another way to reduce the size.
Generally used pooling layers are average and maximum pool layers. 
They are applied similarly to a convolution but capture the average or the maximum of a window.
Neither of these approaches requires any extra learnable weights.

\subsection{Embedding}
Generally, the role of the convolutional layers in a CNN model is to act as a feature extractor.
As can be seen in the number of parameters of the fully connected layer on a full-sized image, the image needs to have a smaller dense representation in order to connect a linear layer to it.
This representation can be generated with the convolutional layers.
The final representation that can be pushed into a linear classifier for some final inference is called an image embedding.
Embeddings are representations of things in relatively small continuous numbered vectors, here we consider embeddings of images, but embeddings are used, for example, for words and sentences as well.
Besides just doing classification, the embeddings can represent things, for example, faces or objects.
When a general embedding is generated it can be used to for example to evaluate if someone is the same person as the embedding or to find similar images.

\begin{figure}[h!] 
\centering 
\includegraphics[width=0.8\textwidth]{imgs/object_classification_architecture.png}
\caption{Image of an example of a convolutional neural network architecture with backbone (a), image classifier (b) and task head (c) labeled. TODO: IMAGE}
\end{figure}

\subsection{Backbone}
The feature extraction part of a convolutional classifier is called the backbone.
Finding better features is often correlated with more parameters and a larger model, but various architectural decisions can also help.
The largest convolutional models have thousands of layers.
Best models often use as large backbones as is possible within the constraints of the specific deployment.
The backbone is thus responsible for creating the image embedding that is used to produce an output.
Some tasks require feature maps taken from the middle of the backbone instead of only relying on the final layer.

\subsection{Head}
The final layers of a classifier are called the head.
The head is responsible for producing the output, and its exact architecture varies from task to task.
It could be just a single linear layer connecting the image embedding to final class probability distribution or an image segmentation model classifying each pixel in the image.
The head can thus contain whatever layers are needed to get the output, combining batch normalization, dropout, convolutional, recurrent, and linear layers depending on what is needed to solve the task.
An example visualization of how the different parts of a convolutional model are connected is shown in Figure 2.1.
The architecture would vary on a task-by-task basis, and the figure shows a basic image classifier architecture.

>>>>

\chapter{Conclusions}
In this thesis, we have given an overview of fundamental computer vision tasks of image classification and object detection.
We described the importance of the ImageNet data set and some of the most significant image classification architectures based on it.
The importance of these models extend beyond just solving the ImageNet dataset, and we covered how it is possible to apply the trained models in new classification problems with transfer learning.

For object detection, we described how the problem and datasets differ from image classification ones.
We also described the different parts of object detectors and how single and two-stage detectors differ from one another and when one might be preferred over the other.
The object detectors also show how important the ImageNet backbones are, as they are an integral part of the object detection architectures also.
The focal loss is an important loss function used by modern single-stage object detectors in allowing a large number of anchor points for localizing the objects in the images.
Finally, we described one of the current state-of-the-art object detection architectures, the EfficientDet, and how it can be trained on a new problem using transfer learning like in the case of image classification.

Here we have also presented multi-task learning as a generalization of transfer learning.
In a multi-task setting, the goal is to solve multiple tasks using some shared representation. 
It differs from transfer learning in that all tasks are optimized continuously, rather than just serving as a starting point.
Doing transfer learning is an excellent way of obtaining good results, even on little data.
With multi-task training, we can utilize the common model architectures as shared parameters for multiple tasks to allow the model to generalize better and run more efficiently.
This is possible because the model is required to use the same capacity to solve multiple problems that may require different features, some of which can also be useful in other tasks. 
Still, sometimes the multi-task setting can be detrimental to accuracy.
Despite the differences in performance, utilizing shared representations means that there are fewer parameters that the model needs to learn, leading to smaller models and shorter inference times.

We saw multiple examples where multi-task learning played an important part in getting models that are efficient and performant.
One of these was the object detection, where the model tries to classify and localize the objects.
Both soft and hard parameter sharing for multi-task learning were covered, with some specialized architectures taking advantage of the fact that multiple tasks are solved at the same time.
Finally, we found that multiple tasks can serve as auxiliary targets for training to aid the model in finding important features.

A multi-task problem setting modifies the training process and introduces some new parameters that can be tuned when training models.
Task-specific sampling ratio during the data sampling process and giving task-specific loss weights were the two new factors that need to be tuned when different tasks are combined.
These weights need to be tuned appropriately for the model to be able to find the features that work for all tasks.
Beyond just tuning parameters, multi-task models may need some architectural fine-tuning in deciding which tasks should have shared representations and just how much should be shared.

The training process differences due to the multi-task problem setting were evaluated through multiple experiments on combining various datasets into a single model.
With these experiments, we empirically verified some of the assumptions of training the models in a multi-task setting.
We noted that it was difficult to find the very best models as different sampling ratios resulted in very different final models.
As the model gets larger and the number of tasks increases, finding the correct ratios gets ever more difficult.
After adding new tasks, the previous ratios might also need adjusting.
Also, we discussed the difficulties of modifying the existing models to contain our desired output heads.
This was most apparent in the case of object detectors. 
The supplied code is relatively pre-packaged to run on the specified training loop, and modifying these model definitions to add new tasks takes some effort.
We saw some improvements and reductions in the model performance, but in general, the multi-task models were close to the single-task counterparts so that reducing the model size might be a good trade-off.

In the end, what we learned is that the success of multi-task learning some tasks can't be pre-determined currently, and some expensive experimentation is needed to arrive at the best multi-task architecture.  
Thus, multi-task learning can be considered as a trade-off of training time complexity for inference-time benefits in accuracy and speed.
